---
title: "Workshop"
author: "sahar chiha"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the 1st Dataset 

Import **freMTPLfreq** dataset:
```{r}
library(CASdatasets)
data("freMTPLfreq",package = "CASdatasets")
View(freMTPLfreq)
```
```{r}
# Function to check if a variable has unique values
has_unique_values <- function(data, column_name) {
  unique_count <- length(unique(data[[column_name]]))
  total_count <- nrow(data)
  
  if (unique_count == total_count) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

# Test the function
column_to_check <- "PolicyID"
if (has_unique_values(freMTPLfreq, column_to_check)) {
  cat("The variable", column_to_check, "has all unique values.\n")
} else {
  cat("The variable", column_to_check, "does NOT have all unique values.\n")
}
```


## Import the 2nd Databset 
Import **freMTPLsev** dataaset:
```{r}
library(CASdatasets) ### RQ_Chaima : You don't need to reload the library; it's already stored in your environment since the first time you loaded it.
data("freMTPLsev",package = "CASdatasets")
View(freMTPLsev)

```

```{r}

### RQ_Chaima : Do you really need to create the function has_unique_values that you have already used before ?
# Function to check if a variable has unique values
has_unique_values <- function(data, column_name) {
  unique_count <- length(unique(data[[column_name]]))
  total_count <- nrow(data)
  
  if (unique_count == total_count) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}

# Test the function
column_to_check <- "PolicyID"
if (has_unique_values(freMTPLsev, column_to_check)) {
  cat("The variable", column_to_check, "has all unique values.\n")
} else {
  cat("The variable", column_to_check, "does NOT have all unique values.\n")
}
```



```{r}
library(dplyr)
freMTPLsev <- freMTPLsev %>%
  group_by(PolicyID) %>%
  summarise(ClaimAmount = sum(ClaimAmount))
print(freMTPLsev)

```
---> Here, we have summed ClaimAmount for each PolicyID.

## Data merging
```{r}
library(dplyr)
freMTPLfreq$PolicyID = as.integer(freMTPLfreq$PolicyID)
Dataset <- freMTPLfreq %>% left_join(freMTPLsev, by = "PolicyID")
print(Dataset)
```

## Exploring the final Database
```{r}
head(Dataset)
```
**variable types**
```{r}
glimpse(Dataset)
```
## Data cleaning:
**Missing values**
```{r}
# Check for NA values in the entire dataset
is.na(Dataset)

# Count NA values per column
colSums(is.na(Dataset))

# Check rows with complete cases (no NA values)
complete.cases(Dataset)
```
--> Only the variable claimAmmount has missing values (nb missing values = 397779).


```{r}
# Check if all instances where claimNB is 0 also have claimAmount as 0
check_claim <- all(is.na(Dataset$claimAmount[Dataset$claimNB == 0]))

# Output the result
if (check_claim) {
  cat("All instances where claimNb is 0 also have claimAmount as NA.\n")
} else {
  cat("There are instances where claimNb is 0 but claimAmount is NOT NA.\n")
}
```
Logically, if a Policy holder didn't affect any claim (claimNb=0),The claimAmmount should be 0.

```{r}
Dataset <- Dataset %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .)))

# Print updated data
print("Data after replacing NA with 0:")
print(Dataset)
```

**Duplicates**
```{r}
# Check for duplicate rows
has_duplicates <- any(duplicated(Dataset))

# Output the result
if (has_duplicates) {
  cat("The dataset contains duplicate rows.\n")
} else {
  cat("The dataset does NOT contain duplicate rows.\n")
}
```


```{r}
print(Dataset)
```



**Outliers**
Checking if there are outliers in the dataset!
```{r}
library(dplyr)
library(tidyr)
library(tidyselect)
library(ggplot2)

#Exclude PolicyID feature
Dataset <- Dataset %>% dplyr::select(-PolicyID)

# Detect numeric columns
numeric_columns <- Dataset %>% dplyr::select(where(is.numeric))

# Detect factor columns
factor_columns <- Dataset %>%dplyr:: select(where(is.factor) | where(is.character))

# Reshape numeric columns for histograms
if (ncol(numeric_columns) > 0) {
  data_numeric_long <- numeric_columns %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value")
  
  # Plot histograms for numeric variables
  ggplot(data_numeric_long, aes(x = value)) +
    geom_histogram(binwidth = 10, fill = "blue", color = "black") +
    facet_wrap(~ variable, scales = "free") +
    labs(title = "Histograms of Numeric Variables",
         x = "Value",
         y = "Frequency") +
    theme_minimal()
} else {
  cat("No numeric columns found.\n")
}

# Reshape factor columns for bar plots
if (ncol(factor_columns) > 0) {
  data_factor_long <- factor_columns %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value")
  
  # Plot bar plots for factor variables
  ggplot(data_factor_long, aes(x = value)) +
    geom_bar(fill = "pink", color = "black") +
    facet_wrap(~ variable, scales = "free") +
    labs(title = "Bar Plots of Factor Variables",
         x = "Value",
         y = "Frequency") +
    theme_minimal()
} else {
  cat("No factor columns found.\n")
}

```
RQ_Chaima : Could you interpret these plots for ClaimAmount and ClaimNb? Personally, I would challenge the visualization of these two variables to understand their distribution, either by adjusting the plot parameters or by applying bijective transformations to the variables, at least to gain visibility on their shape.


```{r}
print(Dataset)
```

Testing the normality of variables:

RQ_Chaima: Why would you test the normality of the quantitative variables ?
```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Detect numeric variables
numeric_columns <- Dataset %>% dplyr::select(where(is.numeric))

# Function to test normality using Kolmogorov-Smirnov test
test_normality_ks <- function(data) {
  results <- data %>%
    summarise(across(everything(), ~ ks.test(., "pnorm", mean = mean(.), sd = sd(.))$p.value))  # KS test
  return(results)
}

# Test normality for all numeric variables
normality_results <- test_normality_ks(numeric_columns)

# Identify normally and non-normally distributed variables
normally_distributed <- normality_results %>% dplyr::select(where(~ all(. > 0.05)))  # p-value > 0.05
non_normally_distributed <- normality_results %>% dplyr::select(where(~ all(. <= 0.05)))   # p-value <= 0.05

# Print results
cat("Normally Distributed Variables:\n")
print(names(normally_distributed))

cat("\nNon-Normally Distributed Variables:\n")
print(names(non_normally_distributed))

# Visualize distributions using density plots and Q-Q plots
for (col in names(numeric_columns)) {
  # Density plot
  density_plot <- ggplot(Dataset, aes(x = .data[[col]])) +
    geom_density(fill = "blue", alpha = 0.5) +
    labs(title = paste("Density Plot of", col),
         x = col,
         y = "Density") +
    theme_minimal()
  
  # Q-Q plot
  qq_plot <- ggplot(Dataset, aes(sample = .data[[col]])) +
    geom_qq() +
    geom_qq_line(color = "red") +
    labs(title = paste("Q-Q Plot of", col),
         x = "Theoretical Quantiles",
         y = "Sample Quantiles") +
    theme_minimal()
  
  # Display plots
  print(density_plot)
  print(qq_plot)
}
```

----> all the numeric variables are not normally distributed. Therefore, the Interquartile Range (IQR) Method is proposed to verify the existence of the outliers.



```{r}
# Load necessary libraries
library(dplyr)

# Function to check for outliers using the IQR method
check_outliers_iqr <- function(data, column) {
  # Calculate Q1, Q3, and IQR
  Q1 <- quantile(data[[column]], 0.25, na.rm = TRUE)
  Q3 <- quantile(data[[column]], 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  
  # Define lower and upper bounds for outliers
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  
  # Detect outliers
  outliers <- data[[column]] < lower_bound | data[[column]] > upper_bound
  
  # Return TRUE if outliers are present, FALSE otherwise
  return(any(outliers, na.rm = TRUE))
}

# Check for outliers in all numeric columns
numeric_columns <- names(Dataset)[sapply(Dataset, is.numeric)]  # Get numeric column names
outlier_summary <- sapply(numeric_columns, function(col) check_outliers_iqr(Dataset, col))

# Print the outlier summary
print(outlier_summary)

# Optional: Visualize outliers using boxplots
library(ggplot2)

for (col in numeric_columns) {
  boxplot <- ggplot(Dataset, aes(x = .data[[col]])) +
    geom_boxplot(fill = "lightblue") +
    labs(title = paste("Boxplot of", col),
         x = col,
         y = "Value") +
    theme_minimal()
  
  print(boxplot)
}

```

RQ_Chaima : Could you defend the transformations that you chose and explain them?

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(MASS)

# Log transformation for claimAmount
Dataset$ClaimAmount <- log(Dataset$ClaimAmount + 1)

# Square root transformation for claimNb
Dataset$ClaimNb <- sqrt(Dataset$ClaimNb)

# Shift CarAge to make all values positive
Dataset$CarAge <- Dataset$CarAge + 1

# Box-Cox transformation for carAge
bc <- boxcox(lm(CarAge ~ 1, data = Dataset))
lambda <- bc$x[which.max(bc$y)]
Dataset$CarAge <- (Dataset$CarAge^lambda - 1) / lambda

# Cap extreme values in driveAge at the 95th percentile
cap_value <- quantile(Dataset$DriverAge, 0.95, na.rm = TRUE)
Dataset$DriverAge[Dataset$DriverAge > cap_value] <- cap_value

# Standardize density
Dataset$Density <- scale(Dataset$Density)

# Visualize transformed data
ggplot(Dataset, aes(y = ClaimAmount)) +
  geom_boxplot() +
  labs(title = "Boxplot of Transformed claimAmount")
```
## Data description:
```{r}
summary(Dataset)
```


## Analyse univariée:

```{r}
print(Dataset)
```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)


# Identify categorical var:
qualitative_vars <- names(Dataset)[sapply(Dataset, function(x) is.character(x) | is.factor(x))]

# Visualize plots of each categorical var
for (var in qualitative_vars) {
  p <- ggplot(Dataset, aes(x = Dataset[[var]])) +
    geom_bar(fill = "steelblue") +
    labs(
      title = paste("Barplot of", var),
      x = var,
      y = "Frequency"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))  
  
  print(p)
}
```


```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Identify numeric columns
numeric_cols <- sapply(Dataset, is.numeric)

#Extract numeric comlumns
numeric_vars <- Dataset[, numeric_cols, drop = FALSE]

# Descrptive statistic for numeric columns
summary(numeric_vars)

# Histogram for each numeric var
for (col in colnames(numeric_vars)) {
  hist(numeric_vars[[col]], 
       main = paste("Histogram of", col), 
       xlab = col, 
       col = "steelblue")
}

# Boxplot for each numerci var
for (col in colnames(numeric_vars)) {
  boxplot(numeric_vars[[col]], 
          main = paste("Boxplot of", col), 
          col = "orange")
}


```



## Analyse bivariée:
```{r}
print(Dataset)
```

```{r}
library(corrplot)

cor_matrix <- cor(Dataset[sapply(Dataset,is.numeric)], method = "pearson")
cor_matrix
```
**Correlation Matrix:**
```{r}
cor_matrix <- corrplot(cor_matrix, method = "color", type = "upper", tl.col = "black", tl.srt = 45)
```

 
##Analyse bivariée:

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(stringi)

# Load the French regions map
url <- "https://github.com/gregoiredavid/france-geojson/raw/master/regions.geojson"
France_map <- st_read(url)


# Clean and harmonize region names in your dataset
Dataset <- Dataset %>%
  mutate(Region = tolower(trimws(Region)),  # Clean and lowercase region names
         Region = stri_trans_general(Region, "Latin-ASCII"))  # Remove accents

# Create a mapping of old region names to new region names
region_mapping <- c(
  "aquitaine" = "nouvelle-aquitaine",
  "basse-normandie" = "normandie",
  "haute-normandie" = "normandie",
  "centre" = "centre-val de loire",
  "limousin" = "nouvelle-aquitaine",
  "nord-pas-de-calais" = "hauts-de-france",
  "pays-de-la-loire" = "pays de la loire",
  "poitou-charentes" = "nouvelle-aquitaine"
)

# Apply the mapping to your dataset
Dataset <- Dataset %>%
  mutate(Region = ifelse(Region %in% names(region_mapping), region_mapping[Region], Region))

# Aggregate data by region
region_claims <- Dataset %>%
  group_by(Region) %>%
  summarise(Total_Claim = sum(ClaimNb, na.rm = TRUE),
            Total_ClaimAmount = sum(ClaimAmount, na.rm = TRUE))

# Clean and harmonize region names in France_map
France_map <- France_map %>%
  mutate(nom = tolower(trimws(nom)),
         nom = stri_trans_general(nom, "Latin-ASCII"))

# Merge your dataset with France_map
France_map <- France_map %>%
  left_join(region_claims, by = c("nom" = "Region"))

# Visualize the number of claims
ggplot(data = France_map) +
  geom_sf(aes(fill = Total_Claim), color = "black", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "red", na.value = "grey90", name = "Nb Sinistres") +
  labs(title = "Nombre de sinistres par région en France") +
  geom_sf_text(aes(label = nom), color = "black", size = 3) +
  theme_minimal()

# Visualize the total claim amount
ggplot(data = France_map) +
  geom_sf(aes(fill = Total_ClaimAmount), color = "black", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "red", na.value = "grey90", name = "Montant de Sinistres") +
  labs(title = "Montant de sinistres par région en France") +
  geom_sf_text(aes(label = nom), color = "black", size = 3) +
  theme_minimal()
```
## Discretization of CarAge and DriverAge:
```{r}
 # Load necessary libraries
library(dplyr)

# Discretization of DriverAge into classes
Dataset <- Dataset %>%
  mutate(
    DriverAge_Class = cut(
      DriverAge,
      breaks = c(0, 25, 40, 60, Inf),  # Define the intervals
      labels = c("Young", "Adult", "Senior", "Elderly"),  # Class names in English
      right = FALSE  # Include the lower bound, exclude the upper bound
    )
  )

# Discretization of CarAge into classes
Dataset <- Dataset %>%
  mutate(
    CarAge_Class = cut(
      CarAge,
      breaks = c(0, 5, 10, 15, Inf),  # Define the intervals
      labels = c("New", "Young", "Middle-aged", "Old"),  # Class names in English
      right = FALSE  # Include the lower bound, exclude the upper bound
    )
  )

# Display the dataset with the new discretized variables
print(Dataset)
 
      

```
```{r}

# Calculate average claim amount for each Brand and DriverAge
average_claim <- Dataset %>%
  group_by(Brand, DriverAge) %>%  # Group by Brand and AgeClass
  summarise(
    AverageClaimAmount = mean(ClaimAmount, na.rm = TRUE)  # Calculate average claim amount
  ) %>%
  ungroup()  # Ungroup to return a regular dataframe

# Display the result
print(average_claim)

```


```{r}
   # Load ggplot2 for visualization
library(ggplot2)

# Plot for DriverAge classes
ggplot(Dataset, aes(x = DriverAge_Class)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Driver Age Classes", x = "Driver Age Class", y = "Frequency") +
  theme_minimal()

# Plot for CarAge classes
ggplot(Dataset, aes(x = CarAge_Class)) +
  geom_bar(fill = "orange") +
  labs(title = "Distribution of Car Age Classes", x = "Car Age Class", y = "Frequency") +
  theme_minimal()
```
##Glm:
  **TARGET FEATURE: ClaimAmmount **
##Split the Data into Training and Test Sets
```{r}
# Créer un index pour diviser les données
train_index <- sample(1:nrow(Dataset), 0.7 * nrow(Dataset))

# Diviser les données
train_data <- Dataset[train_index, ]  # 70% pour l'entraînement
test_data <- Dataset[-train_index, ]  # 30% pour le test

# Vérifier les dimensions des ensembles
print(paste("Taille de l'ensemble d'entraînement :", nrow(train_data)))
print(paste("Taille de l'ensemble de test :", nrow(test_data)))
```

##Check the Distribution of the Target Variable
```{r}
# Check the distribution of the target variable
summary(train_data$ClaimAmount)
summary(test_data$ClaimAmount)

# Visualize the distribution (optional)
par(mfrow = c(1, 2))
hist(train_data$ClaimAmount, main = "Distribution of ClaimAmount (Training)", xlab = "ClaimAmount")
hist(test_data$ClaimAmount, main = "Distribution of ClaimAmount (Test)", xlab = "ClaimAmount")
```

##Fit the GLM Model on the Training Set

RQ_Chaima : Based on the histogram of ClaimAmount, do you think that the Gaussian family is the right one to model the losses? Personally, I notice a distribution with a mass at 0, asymmetric to the left, which is totally different from a Gaussian distribution, which is characterized by a symmetric shape.

In addition to that, why would we use both the continuous and categorized variables of DriverAge, which are surely dependent ? The same applies to CarAge. 

Do you think that we can use ClaimNB to predict ClaimAmount? I suppose that it's a random variable that we need to predict, right? The insurer doesn’t have this information at the time of underwriting the contract.

```{r}
# Fit the initial GLM model with all variables
model <- glm(ClaimAmount ~ ., data = train_data, family = gaussian(link = "identity"), offset = log(train_data$Exposure))

# Display the model summary
summary(model)
```


##Evaluate the Model on the Test Set
```{r}

# Check for duplicates
print(sum(duplicated(test_data)))

# Remove duplicates (if necessary)
test_data <- unique(test_data)

# Predict values on the test set
predictions <- predict(model, newdata = test_data, type = "response")

# Check lengths
print(length(test_data$ClaimAmount))
print(length(predictions))

# Calculate the error (e.g., Mean Squared Error)
mse <- mean((test_data$ClaimAmount - predictions)^2)
print(paste("Mean Squared Error (MSE) on the test set :", mse))

# Visualize predictions vs actual values
plot(test_data$ClaimAmount, predictions, main = "Predictions vs Actual", xlab = "Actual Values", ylab = "Predictions")
abline(0, 1, col = "red")
```
```{r}
regroup_modalites <- function(model, data) {
  summary_model <- summary(model)
  
  # Ensure the coefficients matrix is valid
  if (!is.matrix(summary_model$coefficients)) {
    stop("Unexpected format: coefficients are not in matrix format.")
  }

  p_values <- summary_model$coefficients[, "Pr(>|t|)"]
  
  # Identify non-significant levels
  non_significatives <- which(p_values > 0.05)

  if (length(non_significatives) == 0) {
    message("No non-significant levels found.")
    return(data)
  }

  for (i in non_significatives) {
    var_name <- rownames(summary_model$coefficients)[i]
    
    if (grepl("\\(", var_name)) {  # Check if it's a categorical level
      base_var <- sub("\\(.*", "", var_name)
      modalite <- sub(".*\\((.*)\\).*", "\\1", var_name)
      
      if (base_var %in% names(data)) {
        data[[base_var]] <- factor(replace(as.character(data[[base_var]]), 
                                           data[[base_var]] == modalite, 
                                           "Reference"))
      } else {
        warning(paste("Variable", base_var, "not found in data."))
      }
    }
  }

  return(data)
}

# Apply the grouping function to the training set
train_data <- regroup_modalites(model, train_data)

   
```


```{r}
# Apply the stepwise method
step_model <- step(model, direction = "both")

# Display the final model summary
summary(step_model)
```



